Identifying Departures from the Fully Developed Speckle Hypothesis in Intensity SAR Data with Non-Parametric Estimation of the Entropy\\ \thanks{Grantee Capes.}

### 2023-10-16 Text/Articles/MIGARS_2024/migars2024_abstract.tex
report and code

\documentclass[conference]{IEEEtran}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{booktabs}                        %
\usepackage[utf8]{inputenc}                  % 
\usepackage{rotating}                        %
%\usepackage{subfigure}                       
%\usepackage[export]{adjustbox}              
\usepackage{bm}

@@

\DeclareMathOperator{\traco}{tr}

@@

In this paper we consider an estimator based on spacings.

@@

\section{Statistical modeling for SAR data}\label{sec_02}

\subsection{Intensity data}

The primary models used for intensity SAR data include the Gamma, $\mathcal{K}$, and $\mathcal{G}^0$ distributions.

\begin{align}
	f(z; \mu, L)&=\frac{L^L}{\Gamma(L)\mu(L)}z^{L-1}\exp\left\{-Lz/\mu\right\},\\
	f(z; \alpha, \gamma, L)&=\frac{L^L\Gamma(L-\alpha)}{\gamma^{\alpha}\Gamma(-\alpha)\Gamma(L)}z^{L-1}(\gamma+Lz)^{\alpha-L}
\end{align}

\section{Nonparametric entropy estimation}

Suppose a random variable $X$ has a distribution function $F(x)$ with a continuous density function $f(x)$. Then, the Shannon entropy $H(X)$ is defined as
\begin{equation*}
  %\label{E:entropy2}
  H(X)=-\int_{-\infty }^\infty \ f(x)\log(f(x))\,dx.
\end{equation*}
The problem of estimating $H(X)$ has been considered by many authors including \cite{vasicek1976test, Bert1992, Wieczorkowski1999, correa1995new}, who proposed estimators based on spacings.

Vasicek \cite{vasicek1976test}, used $f(x)=p$  to express $ H(f)$ as
\begin{equation*}
	H(X)= \int_0^1 \log\left(\frac{d}{dp}Q(p)\right)dp,
\end{equation*}
where $Q(p)=F^{-1}(p)=\inf\left\{x: F(x)\leq p\right\}$ is the quantile function. The derivative of $F^{-1}(p)$ is  estimated by a function of the order statistics.
Assuming that  $\bm{X}=(X_1, X_2, \ldots,X_n)$ is a random sample from the distribution $F(x)$, the estimator is given by
\begin{equation*}
%\label{E:VanEs}
	\widehat{H}_{V}(\bm{X})=\frac{1}{n}\sum_{i=1}^{n}\log\left[\frac{n}{2m}\left(X_{(i+m)}-X_{(i-m)}\right)\right],
	\end{equation*}
where $m<n/2$ is a positive integer, $X_{(i+m)}-X_{(i-m)}$ is the $m$-spacing and $X_{(1)}\leq X_{(2)}\leq\ldots\leq X_{(n)}$ are the order statistics and $X_{(i)}= X_{(1)}$ if $i<1$, $X_{(i)}= X_{(n)}$ if $i>n$.

Multiple authors have presented  adaptations to Vasicek's estimator, including Van Es \cite{Bert1992}, who proposed a new estimator of entropy given by
\begin{multline}
%\label{E:VanEs}
	\widehat{H}_{VE}(\bm{X})=\frac{1}{n-m}\sum_{i=1}^{n-m}\log\left[\frac{n+1}{m}\left(X_{(i+m)}-X_{(i)}\right)\right]\\
	+\sum_{k=m}^n\frac{1}{k}+\log\frac{m}{n+1}.
\end{multline}
Van Es demonstrated that, under some conditions, this estimator exhibits consistency and asymptotic normality.

\section{Fully Developed Speckle}

@@

\section{Conclusion}\label{sec_09}

>>>>

\documentclass[conference]{IEEEtran}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{booktabs}  
\usepackage{enumitem}
\usepackage{float}
                      %
\usepackage[utf8]{inputenc}                  % 
\usepackage{rotating}                        %
%\usepackage{subfigure}                       
%\usepackage[export]{adjustbox}              
\usepackage{bm}

@@

\DeclareMathOperator{\traco}{tr}
%\theoremstyle{plain}
\newtheorem{lemma}{Lemma}

@@

%In this paper we consider an estimator based on spacings.

@@

\section{BACKGROUND}\label{sec_02}

\subsection{Statistical modeling for Intensity SAR data}

The primary models used for intensity SAR data include the Gamma and $\mathcal{G}_I^0$  distributions~\cite{Frery1997}. 
The first is suitable for fully developed speckle and can also be considered a limiting case within the $\mathcal{G}_I^0$ distribution.
Whereas the latter is appealing due to its versatility in accurately representing regions with various roughness characteristics~\cite{Cassetti2022}.
We denote $Z \sim \Gamma_{\text{SAR}}(L, \mu)$ and $Z \sim G_I^0(\alpha, \gamma, L)$ to indicate that $Z$ follows the distributions characterized by the respective probability density functions:

\begin{align}
	f_Z(z;L, \mu)&=\frac{L^L}{\Gamma(L)\mu^L}z^{L-1}\exp\left\{-Lz/\mu\right\},\label{E:gamma1}\\
	f_Z(z; \alpha, \gamma, L)&=\frac{L^L\Gamma(L-\alpha)}{\gamma^{\alpha}\Gamma(-\alpha)\Gamma(L)}\cdot\frac{z^{L-1}}{(\gamma+Lz)^{L-\alpha}},\label{E:gi01}
\end{align}
where, in~\eqref{E:gamma1} $\mu > 0$ is the mean; in~\eqref{E:gi01}  $\gamma > 0$ is the scale, $\alpha < -1$ measures the roughness,  $L \geq 1$ is the number of looks, and $\Gamma(\cdot)$ is the gamma function.

From \eqref{E:gi01}, the $r$th moment of $Z$ is expressed as:
\begin{align}
	E_{G_I^0}\left(Z^r\right)=\left(\frac{\gamma}{L}\right)^r\frac{\Gamma(-\alpha-r)}{\Gamma(-\alpha)}\cdot\frac{\Gamma(L+r)}{\gamma(L)}, \quad \alpha <-r. 
	\label{E:rmom}
\end{align}

Even though the $\mathcal{G}_I^0$  distribution is defined by the parameters $\alpha$ and $\gamma$, SAR literature commonly utilizes the texture $\alpha$ and the mean $\mu$~\cite{Nascimento2010}.
In this way, we compute the expected value $\mu$ using the expression in~\eqref{E:rmom}, and we reparametrize~\eqref{E:gi01} using $\mu$, $\alpha$, and $L$. Then
\begin{align*}
	\mu=\left(\frac{\gamma}{L}\right)\frac{\Gamma(-\alpha-1)}{\Gamma(-\alpha)}\cdot\frac{\Gamma(L+1)}{\gamma(L)}=-\frac{\gamma}{\alpha+1}.
\end{align*}
Thus, the probability density functions that characterize the $G_I^0(\mu, \alpha, L)$ is
{\small % o \footnotesize
\begin{align}
		f_Z(z; \mu, \alpha, L)=\frac{L^L\Gamma(L-\alpha)}{(-\mu(\alpha+1))^{\alpha}\Gamma(-\alpha)\Gamma(L)}\frac{z^{L-1}}{(-\mu(\alpha+1)+Lz)^{L-\alpha}}.\label{E:gi02}
\end{align}
}

\subsection{Parametric \& Non-Parametric Shannon Entropy}

The parametric representation of Shannon entropy for a system described by a continuous random variable is as follows:
\begin{equation}
  \label{E:entropy2}
  H_(Z)=-\int_{-\infty }^\infty \ f(z)\log(f(z))\, \mathrm{d}z,
\end{equation}
here, $f(\cdot)$ is the probability density function that characterizes the distribution of the real-valued random variable $Z$.

Using~\eqref{E:entropy2}, we can express the Shannon entropy of $\Gamma_{\text{SAR}}$ in~\eqref{E:gamma1} and $\mathcal{G}_I^0$ in~\eqref{E:gi02} based on~\cite{Cassetti2022, A.Ferreira2020}:
%{\small % o \footnotesize
\begin{align}
H_{\Gamma_{\text{SAR}}}(L, \mu) &=   L -\log(L/\mu)+\log\Gamma(L)+(1-L)\psi^{(0)}(L), \label{E:E-gamma}\\
H_{\mathcal{G}_I^0}(\mu, \alpha, L) &= -\log(\alpha)+\log(\mu)+\log(1+\alpha)\nonumber \\
&\quad -(1-\alpha)\psi^{(0)}(-\alpha) +\log(-\alpha)-\log(L) \nonumber \\
&\quad + (L-\alpha)\psi^{(0)}(L-\alpha) +\log\left(\mathcal{B}(L,-\alpha)\right)\nonumber \\
&\quad  + (1-L)\psi^{(0)}(L), \label{E:E-GIO}
\end{align}
where $\psi^{(0)}(\cdot)$ and $\mathcal{B}(\cdot)$ are the digamma and beta functions, respectively.

%-------------------------------------------------------------------------------------------------------------

%\section{Nonparametric entropy estimation}

The problem of non-parametric estimating of $H(Z)$ has been considered by many authors including~\cite{vasicek1976test, Bert1992, Wieczorkowski1999, correa1995new}, who proposed estimators based on spacings.

Vasicek \cite{vasicek1976test}, used $f(z)=p$  to express $ H(Z)$ as
\begin{equation*}
	H(Z)= \int_0^1 \log\frac{\mathrm{d}}{\mathrm{d}p}Q(p)\mathrm{d}p,
\end{equation*}
where $Q(p)=F^{-1}(p)=\inf\left\{z: F(z)\leq p\right\}$ is the quantile function. The derivative of $F^{-1}(p)$ is  estimated by a function of the order statistics~\cite{AlOmari2019}.

Assuming that  $\bm{Z}=(Z_1, Z_2, \ldots,Z_n)$ is a random sample from the distribution $F(z)$, the estimator is defined as:
\begin{equation*}
%\label{E:VanEs}
	\widehat{H}_{V}(\bm{Z})=\frac{1}{n}\sum_{i=1}^{n}\log\left[\frac{n}{2m}\left(Z_{(i+m)}-Z_{(i-m)}\right)\right],
	\end{equation*}
where $m<n/2$ is a positive integer, $Z_{(i+m)}-Z_{(i-m)}$ is the $m$-spacing and $Z_{(1)}\leq Z_{(2)}\leq\ldots\leq Z_{(n)}$ are the order statistics and $Z_{(i)}= Z_{(1)}$ if $i<1$, $Z_{(i)}= Z_{(n)}$ if $i>n$.

Multiple authors have presented  adaptations to Vasicek's estimator, including Van Es~\cite{Bert1992}, who proposed a new estimator of entropy given by:
\begin{multline}
\label{E:VanEs}
	\widehat{H}_{\text{VE}}(\bm{Z})=\frac{1}{n-m}\sum_{i=1}^{n-m}\log\left[\frac{n+1}{m}\left(Z_{(i+m)}-Z_{(i)}\right)\right]\\
	+\sum_{k=m}^n\frac{1}{k}+\log\frac{m}{n+1}.
\end{multline}

Van Es proved that, under general conditions,~\eqref{E:VanEs} converges almost surely to $H(Z)$ when $m, n\rightarrow\infty$,  $m/ \log(n)\rightarrow\infty$, and $m/n \rightarrow0$.
The author also proved the asymptotic normality of the estimator when $m, n\rightarrow\infty$ and $m=o(n^{1/2})$.
\section{Hypothesis testing based on non-parametric entropy}

General asymptotic results for functions of spacings are detailed in~\cite{Khashimov1990}, while Van Es~\cite{Bert1992} has developed a correction for the case of Shannon entropy.
Following the work of these authors, the next result applies:

\begin{lemma}
Suppose that $f(\cdot)$ is a bounded density bounded away from zero and satisfies a Lipschitz condition on its support.
Then the next asymptotic results follow: If $m,n\rightarrow \infty$ and $m=o(n^{1/2})$, then:
$$
\sqrt{n}\,\Big(\widehat{H}_{\text{VE}}+\int_{-\infty}^\infty f(z)\log f(z) \mathrm{d}z\Big)
\xrightarrow[]{\mathcal{D}}
\mathcal{N}(0,\operatorname{Var}(\log f(Z))),
$$
where $o(\cdot)$ represents the little-o and $\widehat{H}_{\text{VE}}$ is defined in expression~\eqref{E:VanEs}.

\end{lemma}

Let us consider, starting from the previous lemma, the test of the null hypothesis $\mathcal{H}_0: T_{\mathcal{D}}=D_0$ as opposed to one of the other three:
%\begin{enumerate}[label=(\roman*)]
  %\item $\mathcal{H}_1:T_{\mathcal{D}}\neq D_0,$
  %\item $\mathcal{H}_1:T_{\mathcal{D}}> D_0,$
  %\item $\mathcal{H}_1:T_{\mathcal{D}}< D_0.$
%\end{enumerate}
$$
\text{(i)}\, \mathcal{H}_1:T_{\mathcal{D}}\neq D_0, \, \, \text{(ii)}\, \mathcal{H}_1:T_{\mathcal{D}}> D_0, \, \, \text{or}\, \,  \text{(iii)}\, \mathcal{H}_1:T_{\mathcal{D}}< D_0. 
$$

For this purpose, we can use the test statistics:
$$
Z_{m,n} = \frac{\sqrt{n}(\widehat{H}_{\text{VE}}-D_0)}{\sqrt{\operatorname{Var}(\log f(Z))}},
$$
So the null hypothesis should be rejected if (i) $Z_{n,m} > z_{\alpha/2}$ or $Z_{n,m} < - z_{\alpha/2}$ for $\Phi_{\mathcal N}(z_{\alpha/2})=1-\alpha/2$ and $\Phi_{\mathcal N}$ being the standard normal cdf, (ii) $Z_{n,m} > z_{\alpha/2}$ or (iii) $Z_{n,m} < - z_{\alpha/2}$.

The power function for case (i) (two-sided test) at $t\neq D_0$ is given by
\begin{multline*}
\pi_{m,n}(t)=1-\Phi_{m,n}\Big(z_{\alpha/2}-\frac{\sqrt{n}(Z_{m,n}-D_0)}{\sigma}\Big)\\+\Phi_{m,n}\Big(-z_{\alpha/2}-\frac{\sqrt{n}(Z_{m,n}-D_0)}{\sigma}\Big),
\end{multline*}
for a sequence of cdfs $\Phi_{m,n}(x)$ which tends uniformly to $\Phi_{\mathcal N}(z)$.

@@

We verify that the results of the non-parametric entropies for samples of different sizes converge towards the analytical entropies of $\Gamma_{\text{SAR}}(L, \mu)$ and $G^0(\alpha, \mu , L)$ as the sample size increases. 
%This indicates that the data appear to match the fully developed speckle hypothesis.

We assume distinct parameter values for $\mu \in \left\{1, 3, 10\right\}$,  $\alpha\in\left\{-1.5,-3,-5, -8\right\}$, and  $L \in\left\{1,2, 5, 8\right\}$.
We performed experiments to observe convergence under different scenarios, where each employed different sample sizes, i.e., $n\in \left\{9, 25, 49, 81, 121, 500, 5000\right\}$.

For each sample size, we performed $10.000$ samples. Fig. \ref{F4} shows convergence of mean non-parametric entropies, for $\mu=3$ and $L=2$.

%mu <- 3
%L <- 2
%alpha <- -5
%
%r <- 10000
%n_values <- c(9, 25, 49, 81, 121, 500, 1000, 5000, 10000 ) 
\begin{figure}[H] 
\centering
	\includegraphics[scale=0.5]{../../../Figures/PDF/mean_entropies1}\vspace{-1.5mm} 
	\caption{Convergence of mean non-parametric entropies.}
	\label{F4}
\end{figure}
\begin{figure}[H] 
\centering
	\includegraphics[scale=0.5]{../../../Figures/PDF/histo1}\vspace{-1.5mm} 
	\caption{Histogram of entropies.}
	\label{F5}
\end{figure}
\section{Conclusion}\label{sec_09}
